Random number simulation and applications on GPU’s in R?

### 2021-09-03 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
updated document folder extdata folder
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2782 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[article,nojss]{jss}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

@@

@@

\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  University of Toronto\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{ruoyong.xu@mail.utoronto.ca}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
%\SweaveOpts{concordance=FALSE}

@@

\end{leftbar}

@@

\subsection{Creating streams}
In \proglang{clRNG} library, a stream object is a structure that contains three states (see the \proglang{R} output below): the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream. The \proglang{clRNG} function \fct{clrngMrg31k3pCreateStreams} creates streams on the host computer using the MRG31k3p RNG. If we want to use the streams in work items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work item picks streams from the global memory and copies their current states only to its private memory. In \pkg{gpuRandom}, to avoid unnecessary data transfer between host and device, we are able to create streams both on the host computer and on the GPU device. \fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams. \fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on the GPU device. From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used. Below, we show the usage of these two functions by generating 4 streams on the host and on the device respectively.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Create streams on host
streams_on_CPU<-gpuRandom::CreateStreamsCpu(Nstreams=4)
t(streams_on_CPU)
##                 [,1]       [,2]       [,3]       [,4]
## current.g1.1   12345  336690377  502033783  739421137
## current.g1.2   12345  597094797 1322587635 1475938232
## current.g1.3   12345 1245771585 1964121530  730262207
## current.g2.1   12345   85196284 1949818481 1630192198
## current.g2.2   12345  523477687 1607232546  324551134
## current.g2.3   12345 2094976052 1462898381  795289868
## initial.g1.1   12345  336690377  502033783  739421137
## initial.g1.2   12345  597094797 1322587635 1475938232
## initial.g1.3   12345 1245771585 1964121530  730262207
## initial.g2.1   12345   85196284 1949818481 1630192198
## initial.g2.2   12345  523477687 1607232546  324551134
## initial.g2.3   12345 2094976052 1462898381  795289868
## substream.g1.1 12345  336690377  502033783  739421137
## substream.g1.2 12345  597094797 1322587635 1475938232
## substream.g1.3 12345 1245771585 1964121530  730262207
## substream.g2.1 12345   85196284 1949818481 1630192198
## substream.g2.2 12345  523477687 1607232546  324551134
## substream.g2.3 12345 2094976052 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
\fct{gpuRandom::CreateStreamsCpu} has only one argument \code{Nstreams}, which is the number of streams to create. For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) has a default value ``12345''. By setting a seed identical to the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as the streams created on host as follows,
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Create streams on GPU
seed <- c(12345, 12345, 12345, 12345, 12345, 12345)
streams1 <- gpuRandom::CreateStreamsGpu(seedR=seed, Nstreams=4, keepInitial=TRUE)
streams_from_GPU <- as.matrix(streams1)
streams2 = deepcopy(streams1)
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
t(streams_from_GPU)   
##                 [,1]       [,2]       [,3]       [,4]
## current.g1.1   12345  336690377  502033783  739421137
## current.g1.2   12345  597094797 1322587635 1475938232
## current.g1.3   12345 1245771585 1964121530  730262207
## current.g2.1   12345   85196284 1949818481 1630192198
## current.g2.2   12345  523477687 1607232546  324551134
## current.g2.3   12345 2094976052 1462898381  795289868
## initial.g1.1   12345  336690377  502033783  739421137
## initial.g1.2   12345  597094797 1322587635 1475938232
## initial.g1.3   12345 1245771585 1964121530  730262207
## initial.g2.1   12345   85196284 1949818481 1630192198
## initial.g2.2   12345  523477687 1607232546  324551134
## initial.g2.3   12345 2094976052 1462898381  795289868
## substream.g1.1 12345  336690377  502033783  739421137
## substream.g1.2 12345  597094797 1322587635 1475938232
## substream.g1.3 12345 1245771585 1964121530  730262207
## substream.g2.1 12345   85196284 1949818481 1630192198
## substream.g2.2 12345  523477687 1607232546  324551134
## substream.g2.3 12345 2094976052 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
The \fct{gpuRandom::CreateStreamsGpu} has three arguments:
\begin{itemize}
  \item \code{seedR}: an R vector of length 6, which is the seed of streams. Default is set to a vector of six ``12345''.
  \item \code{Nstreams}: number of streams to create.
  \item \code{keepInitial}: logical specifying whether to keep the seed in generated streams. Default is set to TRUE.
\end{itemize}
Streams in \pkg{gpuRandom} are like \fct{.Random.seed} in \proglang{R}. By setting the RNG status with the \fct{set.seed} function in \proglang{R}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs. Similarly in \pkg{gpuRandom}, with the \code{seedR} argument, we are able to produce exactly the same sequence of streams as many times as we want.

@@

In the following code, we use the 4 streams created in Section 2.1 and 4 work items to generate a vector \code{sim_1} of 6 i.i.d. $U (0,1)$ random numbers by the function \fct{gpuRandom::runif}. By default \code{Nglobal} is set as (64,8), and a matrix of \code{prod(Nglobal)} streams created on GPU is supplied. The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host computer.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_1 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
    type = "double")
as.vector(sim_1)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
\end{knitrout}
The arguments of the \fct{gpuRandom::runif} are described as follows:
\begin{itemize}
  \item \code{n}: a number specifying the length if to create a vector, or a vector of length 2 specifying the row and column number if to create a matrix.
  \item \code{streams}: streams used for random number generating.
  \item \code{Nglobal}: global index space that defines the total work-items that execute in parallel. Default is set as $(64,8)$.
  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
\end{itemize}
Next we show that using the same streams (\code{streams1} and \code{streams2}),  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_2 <- gpuRandom::runif(n = 6, streams = streams2, Nglobal = c(1, 4),
    type = "double")
as.vector(sim_2)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
\end{knitrout}
Note that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used, reusing the \code{streams1} will produce a vector  different from \code{sim_1} (result of \code{sim_3}), because each time a random number is generated, the current state of the stream is changed (advances by one position). See below the first six rows (current states) of \code{streams1} after it being used.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
sim_3 = gpuRandom::runif(n = 6, streams = streams1, Nglobal = c(1, 4),
    type = "double")
as.vector(sim_3)
## [1] 0.6487742 0.1112075 0.3661944 0.5018562 0.1088229 0.3114331
t(as.matrix(streams1))[1:6, ]
##            [,1]       [,2]       [,3]       [,4]
## [1,] 1167281028  640923766  502033783  739421137
## [2,] 1918428443 1912157188 1322587635 1475938232
## [3,] 1858462085  286315187 1964121530  730262207
## [4,]  933585541 2119609883 1949818481 1630192198
## [5,] 1132031887  834429287 1607232546  324551134
## [6,]  465230163   47498872 1462898381  795289868
\end{verbatim}
\end{kframe}
\end{knitrout}
In order to reproduce the same sequence of random numbers, we can save the streams in a file on CPU, because GPU device does not keep memory for streams objects after each execution. Below is the \proglang{R} code in which we use the saved streams \code{streams_from_GPU.rds} to produce a vector \code{sim_saved}, which is also same as \code{sim_1}.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
saveRDS(streams_from_GPU, "streams_from_GPU.rds")
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
# Load the streams object as streams_saved
streams_saved <- gpuR::vclMatrix(readRDS("streams_from_GPU.rds"))
sim_saved = gpuRandom::runif(n=6,streams=streams_saved, 
                             Nglobal=c(1,4), type="double")
as.vector(sim_saved)
## [1] 0.7353245 0.5180770 0.6142074 0.2319392 0.1100781 0.3619766
\end{verbatim}
\end{kframe}
\end{knitrout}
We evaluate the distribution of 50000 GPU-generated uniform random numbers by making a density histogram  (the left panel of Figure \ref{fig0}), plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel of Figure \ref{fig0}). The two lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} should come from a standard uniform distribution.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- CreateStreamsGpu(Nstreams =256, keepInitial=1)
random_vector<-gpuRandom::runif(n=50000, streams=streams, 
                                Nglobal=c(64,4), type="double")
random_vector<-as.vector(as.matrix(random_vector))
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}
\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-1} 
\includegraphics[width=0.45\textwidth]{figure/largenumberuniforms-2} 
\end{knitrout}
\caption{\label{fig0} histogram and empirical cumulative distribution of the GPU-generated random numbers.}
\end{figure}

@@

We show the usage of \fct{gpuRandom::rnorm} by generating a $512 \times 1024$ matrix named \code{normal_matrix} of double presicion Gaussian random numbers, and assess Normality by making a histogram (left panel of Figure \ref{fig2}) and a Q-Q plot (right panel of Figure \ref{fig2}) of the generated random numbers. The points in the Q-Q plot in Figure \ref{fig2} fall along the diagonal line, we have good evidence the data are from Gaussian distribution. We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part (adapted from \fct{stat::qnorm}) on GPU, which speeds up the Q-Q plot greatly.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                Nglobal=c(512,128), type="double")
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
avector<-as.vector(as.matrix(normal_matrix))
\end{verbatim}
\end{kframe}
\end{knitrout}
\begin{figure}[H]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
hist(avector,breaks=40, main = "histogram")
gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
\end{verbatim}
\end{kframe}
\includegraphics[width=0.45\textwidth]{figure/histAndQQ-1} 
\includegraphics[width=0.45\textwidth]{figure/histAndQQ-2} 
\end{knitrout}
\caption{\label{fig2} Histogram and Q-Q plot of the GPU-generated random numbers.}
\end{figure}
Let's compare the time of generating a large-size matrix of more than 26.8 million Gaussian random numbers of double precision using the \fct{stats::rnorm} and using the \fct{gpuRandom::rnorm}. Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $128 \times 128$ work-items is about 10 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size goes larger.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
system.time(gpuRandom::rnorm(c(16384,16384), streams=streams, 
                             Nglobal=c(512,128), type="double"))
##    user  system elapsed 
##   0.112   0.001   1.279
system.time(matrix(stats::rnorm(16384*16384),16384,16384))
##    user  system elapsed 
##  11.962   0.898  12.838
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\subsection{Comparing running time: Month data example}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{table}[H]

\caption{\label{tab:monthdata}Month\label{tab:month}}
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
  & Ane & Men & Cya & Her & Omp & Gas & Lim & Cle & Pal & Down & Chro & Hypo\\
\hline
Jan & 29 & 55 & 172 & 46 & 39 & 73 & 48 & 183 & 77 & 103 & 102 & 174\\
\hline
Feb & 25 & 45 & 175 & 35 & 31 & 55 & 34 & 142 & 81 & 115 & 100 & 180\\
\hline
Mar & 31 & 48 & 182 & 41 & 47 & 72 & 40 & 200 & 86 & 90 & 96 & 180\\
\hline
Apr & 34 & 45 & 186 & 36 & 32 & 75 & 42 & 173 & 56 & 87 & 90 & 193\\
\hline
May & 33 & 40 & 187 & 46 & 24 & 80 & 35 & 180 & 75 & 91 & 100 & 197\\
\hline
Jun & 34 & 48 & 189 & 35 & 33 & 75 & 45 & 154 & 74 & 102 & 100 & 182\\
\hline
Jul & 26 & 43 & 198 & 34 & 21 & 74 & 36 & 179 & 79 & 86 & 92 & 193\\
\hline
Aug & 24 & 41 & 189 & 44 & 43 & 62 & 48 & 183 & 88 & 109 & 94 & 194\\
\hline
Sept & 34 & 44 & 147 & 40 & 37 & 66 & 36 & 158 & 73 & 112 & 103 & 196\\
\hline
Oct & 25 & 43 & 207 & 45 & 31 & 65 & 49 & 181 & 77 & 108 & 115 & 220\\
\hline
Nov & 36 & 55 & 188 & 39 & 39 & 62 & 43 & 144 & 68 & 98 & 79 & 173\\
\hline
Dec & 23 & 48 & 196 & 31 & 31 & 71 & 31 & 177 & 86 & 86 & 73 & 156\\
\hline
\end{tabular}
\end{table}

>>>>

\documentclass[article,nojss]{jss}

@@

<<echo=FALSE,results="hide">>=
options(width=65)
@

@@

<<preliminaries, echo=FALSE, results="hide">>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
knitr::opts_chunk$set(highlight=FALSE, background='#FFFFFF', fig.height=4, fig.width=6, out.width='0.45\\textwidth')
library("MASS")
@

@@

\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  University of Toronto\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{ruoyong.xu@mail.utoronto.ca}}

\begin{document}
%\SweaveOpts{concordance=FALSE}

@@

\end{leftbar}
<<echo=FALSE, results="hide", message=FALSE>>=
library("gpuR")
library("gpuRandom")
setContext(   grep('gpu', listContexts()$device_type) [1]    )
setwd("/home/ruoyong/paper1_2021_5")
@

@@

\subsection{Creating streams}
In \proglang{clRNG} library, a stream object is a structure that contains three states (see the \proglang{R} output below): the current state of the stream, the initial state of the stream (or seed), and the initial state of the current substream. The \proglang{clRNG} function \fct{clrngMrg31k3pCreateStreams} creates streams on the host computer using the MRG31k3p RNG. If we want to use the streams in work items on a GPU device, the streams have to be copied from the host to the global memory of the GPU device, and then each work item picks streams from the global memory and copies their current states only to its private memory. In \pkg{gpuRandom}, to avoid unnecessary data transfer between host and device, we are able to create streams both on the host computer and on the GPU device. \fct{gpuRandom::CreateStreamsCpu} is an \proglang{R} interface function of \fct{clrngMrg31k3pCreateStreams} that creates streams on the host as \proglang{R} matrices, so that users can easily view the streams. \fct{gpuRandom::CreateStreamsGpu} is adapted from \fct{clrngMrg31k3pCreateStreams} with the advantage of creating streams directly on the GPU device. From our experience, transferring a large-size matrix of streams from host to device by converting it to a `\code{vclMatrix}' object can take very long time, hence, \fct{gpuRandom::CreateStreamsGpu} is more efficient and more frequently used. Below, we show the usage of these two functions by generating 4 streams on the host and on the device respectively.
<<streams on CPU>>=
# Create streams on host
streams_on_CPU<-gpuRandom::CreateStreamsCpu(Nstreams=4)
t(streams_on_CPU)
@
\fct{gpuRandom::CreateStreamsCpu} has only one argument \code{Nstreams}, which is the number of streams to create. For the MRG31k3p RNG, each state of the stream object is comprised of six 31-bit integers, the state of the first stream (seed of the \proglang{clRNG} library) has a default value ``12345''. By setting a seed identical to the MRG31k3p default seed in \fct{gpuRandom::CreateStreamsGpu}, we can produce on device exactly the same streams as the streams created on host as follows,
<<streams on GPU1, eval=TRUE>>=
# Create streams on GPU
seed <- c(12345, 12345, 12345, 12345, 12345, 12345)
streams1 <- gpuRandom::CreateStreamsGpu(seedR=seed, Nstreams=4, keepInitial=TRUE)
streams_from_GPU <- as.matrix(streams1)
streams2 = deepcopy(streams1)
@ 
<<streams on GPU2, eval=TRUE, echo=FALSE>>=
gpuR::colnames(streams_from_GPU) = c("current.g1.1", "current.g1.2", "current.g1.3", "current.g2.1", "current.g2.2", "current.g2.3",
"initial.g1.1", "initial.g1.2", "initial.g1.3", "initial.g2.1", "initial.g2.2", "initial.g2.3",
"substream.g1.1", "substream.g1.2", "substream.g1.3", "substream.g2.1", "substream.g2.2", "substream.g2.3")   
@ 
<<streams on GPU3,eval=TRUE>>=
t(streams_from_GPU)   
@
The \fct{gpuRandom::CreateStreamsGpu} has three arguments:
\begin{itemize}
  \item \code{seedR}: an R vector of length 6, which is the seed of streams. Default is set to a vector of six ``12345''.
  \item \code{Nstreams}: number of streams to create.
  \item \code{keepInitial}: logical specifying whether to keep the seed in generated streams. Default is set to TRUE.
\end{itemize}
Streams in \pkg{gpuRandom} are like \fct{.Random.seed} in \proglang{R}. By setting the RNG status with the \fct{set.seed} function in \proglang{R}, we are able to repeat the sequence of random seeds, which is an important quality criteria of RNGs. Similarly in \pkg{gpuRandom}, with the \code{seedR} argument, we are able to produce exactly the same sequence of streams as many times as we want.

@@

In the following code, we use the 4 streams created in Section 2.1 and 4 work items to generate a vector \code{sim_1} of 6 i.i.d. $U (0,1)$ random numbers by the function \fct{gpuRandom::runif}. By default \code{Nglobal} is set as (64,8), and a matrix of \code{prod(Nglobal)} streams created on GPU is supplied. The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host computer.
<<eval=TRUE, tidy=TRUE>>=
sim_1 = gpuRandom::runif(n=6,streams=streams1, Nglobal=c(1,4), type="double")
as.vector(sim_1)
@
The arguments of the \fct{gpuRandom::runif} are described as follows:
\begin{itemize}
  \item \code{n}: a number specifying the length if to create a vector, or a vector of length 2 specifying the row and column number if to create a matrix.
  \item \code{streams}: streams used for random number generating.
  \item \code{Nglobal}: global index space that defines the total work-items that execute in parallel. Default is set as $(64,8)$.
  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
\end{itemize}
Next we show that using the same streams (\code{streams1} and \code{streams2}),  \fct{gpuRandom::runif} returns a vector \code{sim_2} of same random numbers as \code{sim_1}. 
<<eval=TRUE,tidy=TRUE>>=
sim_2 <-gpuRandom::runif(n=6, streams=streams2, Nglobal=c(1,4), type = "double")
as.vector(sim_2)
@
Note that \code{streams2} is a \code{deepcopy} of \code{streams1} before \code{streams1} is used, reusing the \code{streams1} will produce a vector  different from \code{sim_1} (result of \code{sim_3}), because each time a random number is generated, the current state of the stream is changed (advances by one position). See below the first six rows (current states) of \code{streams1} after it being used.
<<eval=TRUE,tidy=TRUE>>=
sim_3 = gpuRandom::runif(n=6,streams=streams1, Nglobal=c(1,4), type="double")
as.vector(sim_3)
t(as.matrix(streams1))[1:6,]
@
In order to reproduce the same sequence of random numbers, we can save the streams in a file on CPU, because GPU device does not keep memory for streams objects after each execution. Below is the \proglang{R} code in which we use the saved streams \code{streams_from_GPU.rds} to produce a vector \code{sim_saved}, which is also same as \code{sim_1}.
<<echo=TRUE,eval=FALSE>>=
saveRDS(streams_from_GPU, "streams_from_GPU.rds")
@

<<save streams,eval=TRUE>>=
# Load the streams object as streams_saved
streams_saved <- gpuR::vclMatrix(readRDS("streams_from_GPU.rds"))
sim_saved = gpuRandom::runif(n=6,streams=streams_saved, 
                             Nglobal=c(1,4), type="double")
as.vector(sim_saved)
@
We evaluate the distribution of 50000 GPU-generated uniform random numbers by making a density histogram  (the left panel of Figure \ref{fig0}), plotting its empirical cumulative distribution and overlaying the theoretical cumulative distribution (the right panel of Figure \ref{fig0}). The two lines seem to overlap completely, suggesting the data generated by \fct{gpuRandom::runif} should come from a standard uniform distribution.
<<uniforms_random_vector,eval=TRUE>>=
streams <- CreateStreamsGpu(Nstreams =256, keepInitial=1)
random_vector<-gpuRandom::runif(n=50000, streams=streams, 
                                Nglobal=c(64,4), type="double")
random_vector<-as.vector(as.matrix(random_vector))
@
\begin{figure}[H]
\centering
<<largenumberuniforms, eval=TRUE, echo=FALSE, fig.height=4, fig.width=4, fig.show='hold'>>=
hist(random_vector, prob = TRUE, main="Histogram")
plot(ecdf(random_vector), ylab="F(x)",main="Empirical cumulative distribution")
curve(punif(x, 0, 1), lty=2, add=TRUE, col="red")
@
\caption{\label{fig0} histogram and empirical cumulative distribution of the GPU-generated random numbers.}
\end{figure}

@@

We show the usage of \fct{gpuRandom::rnorm} by generating a $512 \times 1024$ matrix named \code{normal_matrix} of double presicion Gaussian random numbers, and assess Normality by making a histogram (left panel of Figure \ref{fig2}) and a Q-Q plot (right panel of Figure \ref{fig2}) of the generated random numbers. The points in the Q-Q plot in Figure \ref{fig2} fall along the diagonal line, we have good evidence the data are from Gaussian distribution. We find doing a Q-Q plot can occupy a large amount of time in the process of producing an R markdown file, so as an incidental work, we make the function \fct{gpuRandom::qqnorm} that does the quantile computation for the normal distribution part (adapted from \fct{stat::qnorm}) on GPU, which speeds up the Q-Q plot greatly.
<<histAndQQ_prepare,eval=TRUE>>=
streams <- CreateStreamsGpu(Nstreams =512*128, keepInitial=1)
normal_matrix<-gpuRandom::rnorm(c(1024,512), streams=streams, 
                                Nglobal=c(512,128), type="double")
@
<<histAndQQ_prepare_for_cache, eval=TRUE, cache=TRUE>>=
avector<-as.vector(as.matrix(normal_matrix))
@
\begin{figure}[H]
\centering
 <<histAndQQ, echo=TRUE, eval=TRUE, cache=TRUE, fig.height=4, fig.width=4, fig.show='hold'>>=
hist(avector,breaks=40, main = "histogram")
gpuRandom::qqnorm(avector, Nglobal=c(128,64), main="Q-Q Plot")
@
\caption{\label{fig2} Histogram and Q-Q plot of the GPU-generated random numbers.}
\end{figure}
Let's compare the time of generating a large-size matrix of more than 26.8 million Gaussian random numbers of double precision using the \fct{stats::rnorm} and using the \fct{gpuRandom::rnorm}. Both process are not time-consuming, however, using \fct{gpuRandom::rnorm} with $128 \times 128$ work-items is about 10 times (the number may fluctuate a bit from executions) faster than using the \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size goes larger.
<<random normal time compare, eval=TRUE, cache=FALSE>>=
system.time(gpuRandom::rnorm(c(16384,16384), streams=streams, 
                             Nglobal=c(512,128), type="double"))
system.time(matrix(stats::rnorm(16384*16384),16384,16384))
@

@@

\subsection{Comparing running time: Month data example}
<<Timecomparemonth,echo=FALSE,eval=TRUE, results="hide">>=
data2<-read.csv("/home/ruoyong/paper2/month.csv") 
data2<-data2[,-1]
month<-as.matrix(data2)
rownames(month) <- c("Jan", "Feb","Mar","Apr","May","Jun","Jul","Aug","Sept","Oct","Nov","Dec")
colnames(month) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
Nglobal = c(128,128)
streams <- CreateStreamsGpu(Nstreams =128*128, keepInitial=1)
@
<<monthdata,eval=TRUE,echo=FALSE, cache=TRUE>>=
library(knitr)
library(kableExtra)
knitr::kable(head(month, n=12),format="latex",  caption = "Month\\label{tab:month}") %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center",latex_options = "HOLD_position")
@
<<timecomparemonth2Cpu, cache=TRUE, eval=TRUE>>=
#using CPU
stats::fisher.test(month,simulate.p.value = TRUE,B=1e6)$p.value
system.time(stats::fisher.test(month,simulate.p.value = TRUE,B=1e6))
@
\fct{stats::fisher.test} takes about 9.8 seconds (time varies a little bit from executions) for one million simulations, has a p-value of 0.4035.
<<timecomparemonth2Gpu, eval=TRUE,cache=FALSE>>=
#using GPU
almost.1 <- 1 + 64 * .Machine$double.eps
observed_m= -gpuRandom::logfactSum(month, c(16,16))/almost.1
month_GPU<-vclMatrix(month,type="integer")
system.time(gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams, 
            type="double", returnStatistics=FALSE, Nglobal = Nglobal))
@
\fct{gpuRandom::fisher.sim} takes about 2.3 seconds, compared to the elapsed time of \fct{gpuRandom::fisher.sim},  \fct{gpuRandom::fisher.sim} takes less than a quater of the time. 
<<timecomparemonth3Gpu, eval=TRUE,cache=TRUE>>=
result_monthgpu <- gpuRandom::fisher.sim(month_GPU, 1e6, streams=streams,
                   type="double", returnStatistics=TRUE,  Nglobal = Nglobal)
result_monthgpu$simnum
result_monthgpu$counts
result_monthgpu$p.value
@
We requested one million simulations on GPU while the actual number of simulation is 999424, and got 403274 cases whose test statitsic is below the observed threshold, so the p-value from \fct{gpuRandom::fisher.sim} for this table is about 0.4035, which is close enough to the p-value from \fct{stats::fisher.test}: 0.4035.  
<<timecomparemonth4Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
result_monthgpu$cpu <- as.vector(result_monthgpu$sim)
@

\subsection{Comparing running time: Week data example}
<<Timecompareweek, echo=FALSE, eval=TRUE, results="hide">>=
library(latex2exp)
data<-read.csv("/home/ruoyong/paper2/weekday.csv") 
#exampletwo<-as.matrix(as.data.frame(lapply(data, as.numeric)))
data<-data[,-1]
week<-as.matrix(data)
rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Down", "Chro", "Hypo")
@

<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
knitr::kable(head(week, n=7),format="latex",caption = "Week\\label{tab:week}") %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center",latex_options = "HOLD_position")
@
